Course Project Assignment 1

----------------------------------------------

# Loading and preprocessing the data

* Read in activity.csv file, clear the NAs

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
data2 <- data[!is.na(data$steps),]
```

# What is mean total number of steps taken per day?

* Calculate the daily total number of steps, the mean, and median

```{r}
data2$date <- as.character(data2$date)
Date <- levels(as.factor(data2$date))

step.sum <- vector(mode = "numeric", length = length(Date))
step.record <- vector(mode = "numeric", length = length(Date))
step.mean <- vector(mode = "numeric", length = length(Date))
step.median <- vector(mode = "numeric", length = length(Date))

for(i in 1:length(Date)){
    step.sum[i] <- sum(data2$steps[data2$date == Date[i]])
    step.record[i] <- sum(data2$date == Date[i]) 
    step.mean[i] <- step.sum[i]/step.record[i]
    step.median[i] <- median(data2$steps[data2$date == Date[i]])
}
```

* Histogram of total number of steps each day

```{r}
plot1 <- hist(step.sum, breaks = 25, col = "lightblue", main = "Histogram of the Total Number of Steps Each Day", xlab = "Total number of steps each day", ylab = "Frequency", xlim = c(0,25000))
print(plot1)
dev.copy(png, file = "plot1.png")
dev.off()
```

* Summary of daily step records

```{r}
StepSummary <- data.frame(date = Date, StepSum = step.sum, StepRecords = step.record, StepMean = step.mean, StepMedian = step.median)

summary(StepSummary)
```

# What is the average daily activity pattern?

* Average Steps in 5 Minutes Interval Across All Days

```{r}
t <- as.ts(data2$interval[data2$date == Date[1]])

interval.step <- vector(mode = "numeric", length = length(t))
for(i in 1:288){
    interval.step[i] <- sum(data2$steps[data2$interval == t[i]])/53
}

MeanStep <- data.frame(interval = t, steps = interval.step)

peak.time <- t[which(interval.step == max(interval.step))]
  
plot(t, interval.step, type = "l", main = "Average Steps in 5 Minutes Interval Across All Days", col = "blue", xlab ="Time in a day (24 hours)", ylab = "Average steps")

abline(v = peak.time, col = "red")

dev.copy(png, file = "daily-frequency.png")
dev.off()
```

The peak time is at `r peak.time` of the day. 


# Imputing missing values

* Calculate and report the total number of missing values in the dataset

```{r}
na.num <- sum(is.na(data$steps))
```
The total numer of rows with NAs are `r na.num`

* Replace NA values with interval-mean, and forming a new dataset

```{r}
library(dplyr)
na.subset <- data[is.na(data$steps),]
na2 <- merge(na.subset, MeanStep, by = "interval")
Na2Mean <- data.frame(steps = na2$steps.y, date = na2$date, interval = na2$interval)
data3 <- merge(data2, Na2Mean, all = T)
data3 <- arrange(data3, date, interval)
```

* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r}
Date.n <- levels(as.factor(data3$date))

step.sum.n <- vector(mode = "numeric", length = length(Date.n))
step.record.n <- vector(mode = "numeric", length = length(Date.n))
step.mean.n <- vector(mode = "numeric", length = length(Date.n))
step.median.n <- vector(mode = "numeric", length = length(Date.n))

for(i in 1:length(Date.n)){
    step.sum.n[i] <- sum(data3$steps[data3$date == Date.n[i]])
    step.record.n[i] <- sum(data3$date == Date.n[i]) 
    step.mean.n[i] <- step.sum.n[i]/step.record.n[i]
    step.median.n[i] <- median(data3$steps[data3$date == Date.n[i]])
}

plot2 <- hist(step.sum.n, breaks = 25, col = "pink", main = "Histogram of the Total Number of Steps Each Day with NA Replaced", xlab = "Total number of steps each day", ylab = "Frequency", xlim = c(0,25000))
print(plot2)
dev.copy(png, file = "plot2.png")
dev.off()
```
When the NAs in step are replaced by mean steps of that spefic interval, the result is centralization of the histogram (towards the peak).

* Summary of daily step records

```{r}
StepSummary.n <- data.frame(date = Date.n, StepSum = step.sum.n, StepRecords = step.record.n, StepMean = step.mean.n, StepMedian = step.median.n)

summary(StepSummary.n)
```

# Are there differences in activity patterns between weekdays and weekends?

* Make a new factor variable 

```{r}
weekdays <- c("星期一", "星期二", "星期三", "星期四", "星期五")

data3$weekday <- factor(weekdays(as.Date(data3$date)) %in% weekdays, levels = c(TRUE, FALSE), labels = c("weekday", "weekend"))
```

* Calculate the mean step in each group and each interval

```{r}
data.w <- tbl_df(data3)
data.w.mean <- summarize(group_by(data.w, weekday, interval), mean(steps))
colnames(data.w.mean) <- c("weekday", "interval", "mean.step")
```

* Make a facet plot to compare between weekday and weekend

```{r}
library(ggplot2)
g <- ggplot(data.w.mean, aes(x = interval, y = mean.step, color = weekday))
g + geom_line(size = 1) + facet_grid(weekday~.) + labs(x = "Time interval in a day", y = "Mean steps in 5 minutes interval", title = "Mean Steps in Each Time Interval During Weekday and Weekend")

dev.copy(png, file = "compare.png")
dev.off()
```

